Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalogue Data

### 2013-02-27 doc/response.txt

Author's reply to Reviewer's Comments:

Dear Reviewer,

Thankyou for taking the time to review our paper, below are our replies (and a
summary of the actions taken) to each of your points. To aid you in finding
the changes, all of them are shown in bold face in the new manuscript.

1. The paper does not describe the resolution of the convergence maps
from the simulation.  The only number quoted is 30 arcseconds, which
(hopefully!) is not the resolution of the maps.  For example, the
convergence power spectrum peaks on arcminute scales, even neglecting
baryonic effects which only increase small-scale power. Therefore a 30
arcsecond resolution would be woefully inadequate for the purposes of
this study.  The authors should describe their simulations in more
detail, and discuss what fraction of the convergence power is
unresolved by their simulations.

# We have updated section 3.1 to state the resolution of our simulations,
and explain why it should be sufficient for our purposes.

2. My main concern about this work is that the halo occupation
distribution is assumed to be known.  As mentioned in the final
paragraph of the paper, this is a severe limitation which makes it
difficult to assess the validity of the claimed results. In reality,
uncertainties in the HOD will lead to uncertainties in the
reconstruction, and the authors should provide some estimate of these
errors.  For example, the authors could provide an example in which
the "wrong" HOD is assumed, and then compute the resulting errors in
the convergence reconstruction.  One easy example of this could be to
use the stellar masses from the Millennium catalog, which were
discarded in favor of the model of Behroozi et al.

# We agree that this is an important systematic. We have implemented the
suggested test. Swapping Behroozi stellar masses for the Millenium catalogue's
stellar masses, is an extreme change (the catalogue's stellar masses are
unreasonable) but it results in offsets of ~0.015 in kappa for
individual lines of sight. For the ensemble, there is no change in the mean
reconstructed kappa. We have added a discussion of this in section 8.

3. The halo model used in this paper in not described adequately.  For
example, the paper notes that the SHAM model of Behroozi et al. was
adopted for this work, but the paper is overly vague in describing the
implementation.  For example, there is no discussion of how satellites
are treated.  In the absence of any spatial information, a random
galaxy has a significant probability to be either a central galaxy in
a low mass halo, or a satellite in an even lower-mass subhalo inside a
much larger halo.  Is local spatial information ignored in assigning
probabilities for galaxies to be central or satellite?  If so, the
reconstruction will clearly be suboptimal.  For satellites, how are
the subhalo properties assigned?  Their masses and sizes depend on
their location within the host halo.  What properties are assigned to
the host halos of satellites?  Is it assumed that the host halos
(e.g. clusters or groups) have properties determined from their own
central galaxies (e.g. BCGs)?  This may be overly optimistic, as
significant spatial offsets are frequently found between BCG location
and cluster center.  The SHAM model used here must be described in
greater detail.

# We have clarified the details of our halo model in Section 6.1 (new text is shown in bold font). As we now explain, we do not discriminate between satellites and centrals at all.  As you say, this will be sub-optimal, and is one of the areas we are looking to improve upon in further work. However, we do find that it is rare for satellites to contribute a large amount to the total convergence of a line of sight, and have quantified this in the text. Hence, we do not expect this to be a very large source of uncertainty, compared to the choice of M*-Mhalo relation,  and the calibration simulation.

#We neglected to mention the offsets between halo centre and BCG. We have now
mentioned this but it is unlikely to be significant: most of the impact factors
are much greater than typical spatial offsets.

Finally, I have a very minor comment.  In equation (6), it appears
that the mean density of the universe is not subtracted when
estimating kappa_h.  I did not understand why not.

# kappa_h is used as a tracer of kappa_ext, and the integral in equation (7)
converts kappa_h into kappa_ext. Subtracting the mean density of the universe
decreases all values of kappa_h by something akin to the projected mean
density of the universe devided by an effective critical density. Because
subtracting the mean density is a simple transformation in kappa_h the
resulting kappa_ext values do not change (as long as one remembers to
treat the calibration sightlines in the same way as the 'real' sightline).

#Additionally, we have swapped green for black in figures 8 and 9, to make them
clearer in black and white.

We are grateful for the improvements that your comments have made to the paper.
Regards,
Tom Collett (et al.)

### 2013-02-27 doc/paper-lcr.tex
finalized draft
draft update

\author[Collett \etal]{%
  Thomas~E.~Collett$^{1}$\thanks{\collettemail},
  Philip~J.~Marshall$^{2}$,
  Matthew~W.~Auger$^{1}$,
  Stefan~Hilbert$^{3}$,
\newauthor{%
  Sherry~H.~Suyu$^{3,4}$,
  Zachary~Greene$^{4}$,
  Tommaso~Treu$^{4}$\thanks{\packard},
  Christopher~D.~Fassnacht$^{5}$,}
\newauthor{%
  L\'eon~V.~E.~Koopmans$^{6}$,
  Maru\v{s}a Brada\v{c}$^{5}$,
  Roger~D.~Blandford$^{3}$} 
  \medskip\\
  $^1$\ioa\\
  $^2$\oxford\\
  $^3$\kipac\\
  $^4$\ucsb\\
  $^5$\davis\\
  $^6$\kapteyn
}

@@

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight. In
time delay gravitational lenses, ``external convergence,'' $\kappax$, can
dominate the uncertainty in the inferred distances. In this paper we attempt
to reconstruct $\kappax$, due to line of sight structure using a simple halo
model.
%
We use mock catalogues from the Millennium Simulation, and calibrate and 
compare our reconstructed $\pr(\kappax|\data)$ to ray-traced $\kappax$ truth
values:  taking into account realistic uncertainties on redshift and stellar
masses, \tom{we find that the reconstruction of $\kappax$ provides an improvement in
precision of $\sim$50\% over previously published schemes.} 
%
We find that the lowest $\kappax$ lines of sight have the best constrained 
$\pr(\kappax)$. In anticipation of future samples with thousands of lenses,
we find that selecting the third of the systems with the
highest precision $\kappax$ estimates gives a  sub-sample of unbiased time
delay distance measurements \phil{with (on average) just} 1\% uncertainty due
to line of sight external convergence effects. Photometric data alone is
sufficient to pre-select the best-constrained lines of sight, and can be done
before investment in lightcurve monitoring. 
%
\phil{Conversely, we show that somehow selecting lines of sight
with high external shear could, with the reconstruction model presented here,
induce biases of up to 1\% in time delay distance.}  We suggest areas for the
improvement of this general analysis framework (including more sophisticated
treatment of high mass structures) that should allow yet more accurate
cosmological inferences to be made.

@@

Since the mass-sheet degeneracy prevents any estimate of $\kappax$ from the 
strong lens modelling, additional information is required. Weak lensing measurements
can been used for rich lines of sight \citep{NakajimaEtal2009, FadelyEtal2009}, but 
for lines of sight with low galaxy density we must attempt to reconstruct the distribution
of mass along the line of sight. Attempts to reconstruct the mass distribution in strong
lens fields have focused on understanding the external shear which seems to be large in most strong lenses.
Surveys by \citet{Fassnacht+Lubin2002,AugerEtal2007,WilliamsEtal2006,MomchevaEtal2006}
all found groups of galaxies hosting or near to known gravitational lenses.
\citet{WongEtal2011} estimated $\Pr(\gammax)$ given the data of \citet{WilliamsEtal2006}
and \citet{MomchevaEtal2006} but found significant discrepancy between the predicted shear
distribution and the external shear demanded by the strong lens model. 
\citeauthor{WongEtal2011} also found that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear.

@@

Large numerical simulations can also be used to estimate global convergence
distributions. \citet{Holder+Schechter2003} and \citet{DalalEtal2005} carried
out ray-tracing calculations in N-body simulations to estimate the
distribution of external shear values.  \citet{HilbertEtal2009} performed
similar ray-tracing experiments through the Millennium Simulation
\citep{SpringelEtal2005}, generating a predicted $\kappax$ at every position
in a simulated sky. \citet{HilbertEtal2009} found that after removing the
matter on the strong lensing plane \MS lines of sight with strong lenses were
not biased towards high $\kappax$, although selection functions for
discovering samples of strong lenses were not taken into account.

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy over-density, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed over-density towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappax$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappax)$, which was then marginalised over when inferring the time
delay distance in this system. A similar procedure was used in \citet{SuyuEtal2012}.
In a companion paper to this one, Greene et al. (submitted) investigate
improvements to this method by weighting the galaxy counts by galaxy
mass, and perpendicular distance to the line of sight, again using the
\MS mock catalogues and their associated $\kappax$ values to construct
$\Pr(\kappax|\data)$. For the most over-dense lines of sight their
results  show a $\sim$20\% improvement over
number counts alone, but little improvement for less dense ones.

In this paper, we combine the halo model reconstruction approach of
\citeauthor{GunnarssonEtal2006}, \citeauthor{JonssonEtal2010}, 
\citeauthor{WongEtal2011} and others, with the
idea of calibrating to simulations from \citeauthor{SuyuEtal2010}
Taking the $\kappax$ values from
\citeauthor{HilbertEtal2009}'s ray-tracing calculations as the ``truth''
that we have to recover, we use the \MS mock galaxy catalogues first to
{\it calibrate} the reconstruction, and account for unseen mass and
voids, and then to {\it test} the accuracy of the line of sight mass
reconstruction under this assumption. Probabilistically assigning mass
and redshift to every observed galaxy in a given observed field, we
generate Monte Carlo sample line of sight mass distributions, and so
construct the PDF $\Pr(\kappax|\data)$ for that field. We then emulate
the combination of many such PDFs to quantify the residual bias that
would be translated to the global (including cosmological) parameters, 
\phil{if left unaccounted for.}
In doing so, we aim to answer the following questions:

@@

In strong lens systems where the source is time variable, the images do not 
vary simultaneously; the optical path length for each image is different due
to relativistic and geometric effects. The difference in optical path length
causes a time-delay between the light curves of each image.
\citet{FalcoEtal1985} showed that the presence of additional matter (in the
form of a mass-sheet) along the line of sight has no observable effect except
to rescale the time-delays. As such, it is necessary to include $\kappax$ in
the lens modelling, if cosmological parameters are to be estimated accurately
and precisely from observed time delays \citep{SuyuEtal2010}.

If there is external convergence present that is not included in the
lens modelling, then the time delay distance -- inferred assuming $\kappax
= 0$ -- will be $(1-\kappax)$ more than the true value of $\Dt$:
\be 
\label{eq:MassSheet:Dtbias}
\Dt^{\rm{true}}=\frac{\Dt^{{\kappax = 0}}}{1-\kappax}.
\ee

@@

Mass distributions that are physically associated with the lens galaxy will
affect the stellar dynamics of the lens galaxy, so that dynamical observations
can break the internal mass-sheet degeneracy by providing an additional
estimate of the lens galaxy's mass
\citep[e.g.,][]{Koopmans2004,SuyuEtal2010}.  Our aim is to quantify the
lensing effect of massive objects, regardless of their position along the line
of sight. \phil{However, in this paper we restrict ourselves to reconstructing
the line of sight portion of the mass distribution in any given field, which
allows us to work in the weak lensing regime. Treatment of sightlines
containing strong lensing events, and the impact of physically-associated mass
distributions on both the strong lensing and stellar dynamical effects are
postponed to further work.}

@@

\citet{HilbertEtal2009} calculated the lensing convergence using the
``lightcone'' dark matter only  output of the \MS, providing 1.7 degree
square  mock sky maps of this quantity. In this process, the dark matter
density from the simulation was projected onto a set of lens planes at
discrete redshifts, adaptively gridded and smoothed, and then the second
derivatives of the  lensing potential required for the components of the
magnification matrix were computed using a multiple lens plane ray tracing
algorithm. The first order approximation to this algorithm \citep[equation 17
of][]{HilbertEtal2009} gives the total convergence at a given sky
position~$\bmath{\theta}$ as the simple summation of the surface density in
each lens plane, weighted by the inverse critical surface density $\Sigma_{\rm
cr}$ for that plane's redshift:
\be
\kappax(\bmath{\theta}) = \sum_i \frac{\Sigma_i(\bmath{\theta})}
                                     {\Sigma_{\rm cr}(z_i,\zs)}.
\ee
\citet{HilbertEtal2009} found that this approximation is accurate to a few percent or better
in $\kappax$, even on 30 arcsecond scales.

@@

@@

We construct mock photometric galaxy catalogues by selecting all \MS galaxies
within a circular aperture of a given angular radius centred on each randomly
selected line of sight position, with apparent $i$-band magnitude brighter
than a given limit. The parent catalogue, from \citet{HilbertEtal2011},
contains galaxy positions and magnitudes, all of which include lensing effects
(deflections and magnifications). We also have access to the underlying galaxy
halo masses and redshifts, which we will use to calibrate the reconstruction,
and also to explore any sources of bias and scatter. Since the true
convergence is that of only the dark matter, we focus on reconstructing the
dark matter halos alone. Our model for doing this is outlined in the next
section.

@@

Cosmological dark matter simulations have shown that dark matter
halos are well-approximated by NFW profiles 
\citep{NFW}. We assume each halo to have a spherical mass distribution with 
density given by
\be\label{eq:rhonfw}
\rho(r) = 
\frac{\rho_0}{\left(r/r_{s}\right)
\left(1+r/r_{s}\right)^{2}}.
\ee
Here, $r_{s}$ is a characteristic scale 
radius of the cluster, representing the point where
the density slope transitions from $r^{-1}$ to $r^{-3}$. This radius is
related to the virial radius of the halo by $r_{s}~=~r_{200}/c$, where $c$ is
the concentration parameter, which can be estimated from the halo's mass,
using a mass--concentration relation. Typically more massive halos are less
concentrated, but there is some scatter; we use the relation of
\citet{Neto2007} to estimate $c$ from the mass enclosed within $r_{200}$,
which we denote as $M_{200}$. We find our results do not change if the
\citet{MaccioEtal2008} mass--concentration relation is used instead.

@@

Many studies have shown that galaxies have total mass density profiles that
are approximately isothermal in their inner regions
\citep[\eg][]{AugerEtal2010} due to the more concentrated stellar mass
component.
This has been confirmed by galaxy-galaxy weak lensing measurements 
\citep[\eg][]{Mandelbaum,Leauthaud}. We note that the two halo term that
dominates the galaxy-galaxy lensing signal at large radii is explicitly taken
into account in our model, where every galaxy is assigned a dark matter halo.
At large radii an NFW-like profile is probably appropriate, but may not be
correct when a halo is very close to the line of sight.

@@

To estimate a halo's contribution to $\kappah$ we must first know the
halo's mass. For the lines of sight of the \MS, halo masses are
known, but for  real lines of sight the halo masses have to be \infered
from whatever data is available. 
In this work we investigate the use of the empirical stellar mass --
halo mass relation for this purpose. The primary advantage of this
approach is that it is a simple, ``one size fits all'' relation that can
be applied regardless of the stellar mass observed. We discuss
additional observables that could be used to improve the precision of
the halo mass inference in \Sref{sec:discuss} below.

@@

The solution we present here is to calibrate the relationship between the halo
convergence $\kappah$ and the overall convergence $\kappax$ (as would be
calculated by ray tracing through the full density field)  using the
simulations themselves. The halo modelling procedure outlined in
\Sref{sec:model:halos} allows  us to estimate  the convergence due to halos
given observations of  mass, redshift and position,
$\pr(\kappah|\data)$, but what we are interested in is the total
convergence along the line of sight $\pr(\kappax|\data)$. We can obtain
this by considering the expression: 
\begin{equation} 
\Pr(\kappax|\data) = 
   \int \dee\kappah  \Pr(\kappax|\kappah,\data) \Pr(\kappah|\data)
   \label{eq:kappaconv}
\end{equation} 
The first term in the integrand relates the convergence due to model halos,
$\kappah$, to the true convergence, $\kappax$. This conditional distribution
can be constructed from the \MS catalogues, by computing
$\kappah$ from the true halo masses and redshifts 
along each selected line of sight, and then
accumulating $\kappah$ and $\kappax$
along a large number of such sightlines. 
For any given line of sight we can then estimate
$\pr(\kappah|\data)$ as described above, and then multiply it by
$\Pr(\kappax|\kappah)$ and integrate out the intermediate parameter
$\kappah$.

@@

We now put the reconstruction \proceedure outlined above into practice, first
inferring $\kappax$ from $\kappah$ in the case of hypothetical galaxy
catalogues with noiseless redshifts and halo masses. The reason for doing this
is to check the validity of the calibration process, and then also to 
investigate the primary sources of the convergence. It will also provide us
with a measure of the intrinsic uncertainty introduced by our assumption of
all illuminated halos being spherically-symmetric truncated NFW halos with
Neto et al concentration--mass relation.

@@

\Fref{fig:jointkh-k} shows the conditional distribution of $\kappax$ given 
$\kappah$, derived from $10^5$ randomly selected \MS lines of sight
and assuming noise-free halo masses and redshifts. 
\phil{(In this case, $\pr(\kappah|\data)$ is almost a delta function since the only uncertainty is halo
concentration, which has minimal effect on $\pr(\kappah|\data)$.  
)}.
% In this case, the PDF : we can read
% off  $\pr(\kappax|\data)$ just by taking a thin vertical slice through this 
% distribution. 
% Note that the joint PDF plotted in this figure is the product
% of  the conditional distribution that we need for calibration, and also the
% prior PDF for $\kappah$ that we need to ensure that the global distribution of
% $\kappah$ values is defaulted to in the case of poor observational data. 
At fixed $\kappah$ we find in \Fref{fig:jointkh-k} that the scatter in 
$\kappax$
grows with $\kappah$; our reconstruction is better at reproducing under-dense
lines of sight than over-dense lines of sight. The effect of ignoring voids
and the smooth mass component is evident in this plot: $\kappah$ is
significantly higher than $\kappax$, at any given $\kappax$ value. 
\phil{Conditioning on $\kappah$ gives a PDF for
$\kappax$ centred at the correct place, by construction.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\includegraphics[width=\textwidth]{figs/where_is_the_kappa.eps}
\caption[magcut]{Which objects dominate the external convergence for a line of
sight? From left to right, the three figures show the cumulative contribution
to the total convergence from individual halos ($\kappah$) as a function of
1) distance from the line of sight, 2) magnitude and 3) redshift.   In each
panel, the solid line is the median cumulative contribution over a large
sample of lines of sight, while dashed and dot-dashed show the ranges enclosing 68
and 95 percent of the lines of sight respectively.}
\label{fig:where}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We first consider a purely random selection of lines of sight. With the purely
photometric reconstruction, there is no evidence of bias, even after combining
20 lines of sight. \tom{This result is a sanity check: if we had chosen lines
of sight with identical $\kappah^{\rm med}$ rather than $\kappah^{\rm
med}\pm0.003$ randomly selected lines of sight would have zero bias by
construction}. If lines of sight hosting time delay lenses are truely random, 
our method allows $\kappax$ to be reconstructed without inducing a bias (under
the basic assumptions that the universe is like the calibration simulation,
and the correct stellar-mass to halo-mass relation has been used).

>>>>

\author[Collett \etal]{%
  Thomas~E.~Collett$^{1}$\thanks{\collettemail},
  Philip~J.~Marshall$^{2}$,
  Matthew~W.~Auger$^{1}$,
  Stefan~Hilbert$^{3}$,
\newauthor{%
  Sherry~H.~Suyu$^{4,3}$,
  Zachary~Greene$^{4}$,
  Tommaso~Treu$^{4}$\thanks{\packard},
  Christopher~D.~Fassnacht$^{5}$,}
\newauthor{%
  L\'eon~V.~E.~Koopmans$^{6}$,
  Maru\v{s}a Brada\v{c}$^{5}$,
  Roger~D.~Blandford$^{3}$} 
  \medskip\\
  $^1$\ioa\\
  $^2$\oxford\\
  $^3$\kipac\\
  $^4$\ucsb\\
  $^5$\davis\\
  $^6$\kapteyn
}

@@

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight. In
time delay gravitational lenses, ``external convergence,'' $\kappax$, can
dominate the uncertainty in the inferred distances and hence cosmological parameters.
 In this paper we attempt
to reconstruct $\kappax$, due to line of sight structure, using a simple halo
model.
%
We use mock catalogues from the Millennium Simulation, and calibrate and 
compare our reconstructed $\pr(\kappax)$ to ray-traced $\kappax$ ``truth''
values; taking into account realistic uncertainties on redshift and stellar
masses. \tom{We find that the reconstruction of $\kappax$ provides an improvement in
precision of $\sim$50\% over galaxy number counts.} 
%
We find that the lowest-$\kappax$ lines of sight have the best constrained 
$\pr(\kappax)$. In anticipation of future samples with thousands of lenses,
we find that selecting the third of the systems with the
highest precision $\kappax$ estimates gives a  sub-sample of unbiased time
delay distance measurements \phil{with (on average) just} 1\% uncertainty due
to line of sight external convergence effects. Photometric data alone are
sufficient to pre-select the best-constrained lines of sight, and can be done
before investment in lightcurve monitoring. 
%
\phil{Conversely, we show that selecting lines of sight
with high external shear could, with the reconstruction model presented here,
induce biases of up to 1\% in time delay distance.}  We suggest areas for the
improvement of this general analysis framework (including more sophisticated
treatment of high mass structures) that should allow yet more accurate
cosmological inferences to be made.

@@

Since the mass-sheet degeneracy prevents any estimate of $\kappax$ from the 
strong lens modelling, additional information is required. Weak lensing measurements
can be used for rich lines of sight \citep{NakajimaEtal2009, FadelyEtal2009}, but 
for lines of sight with low galaxy density we must attempt to reconstruct the distribution
of mass along the line of sight. Attempts to reconstruct the mass distribution in strong
lens fields have focused on understanding the external shear which seems to be large in most strong lenses.
Surveys by \citet{Fassnacht+Lubin2002,AugerEtal2007,WilliamsEtal2006,MomchevaEtal2006,FassnachtEtal2006}
all found groups of galaxies hosting or near to known gravitational lenses.
Is generally found that lens galaxies, reside in overdense environments,
indistinguishable from those occupied by similarly massive galaxies \citep{Auger2008,TreuEtal2009}
\citet{WongEtal2011} estimated $\Pr(\gammax)$ given the data of \citet{WilliamsEtal2006}
and \citet{MomchevaEtal2006} but found significant discrepancy between the predicted shear
distribution and the external shear demanded by the strong lens model. 
\citeauthor{WongEtal2011} also found that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear.

@@

Large numerical simulations can also be used to estimate global convergence
distributions. \citet{Holder+Schechter2003} and \citet{DalalEtal2005} carried
out ray-tracing calculations in N-body simulations \citep{KauffmannEtal1999,WambsganssEtal2004} to estimate the
distribution of external shear values.  \citet{HilbertEtal2009} performed
similar ray-tracing experiments through the Millennium Simulation
\citep{SpringelEtal2005}, generating a predicted $\kappax$ at every position
in a simulated sky. \citet{HilbertEtal2009} found that after removing the
matter on the strong lensing plane \MS lines of sight with strong lenses were
not biased towards high $\kappax$, although selection functions for
discovering samples of strong lenses were not taken into account.

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy over-density, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed over-density towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappax$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappax)$, which was then marginalised over when inferring the time
delay distance in this system. A similar procedure was used in \citet{SuyuEtal2012}.
In a companion paper to this one, Greene et al. (submitted) investigate
improvements to this method by weighting the galaxy counts by observables
including galaxy luminosity and perpendicular distance to the line of sight, again using the
\MS mock catalogues and their associated $\kappax$ values to construct
$\Pr(\kappax|\data)$. For the most over-dense lines of sight their
results  show a $\sim$20\% improvement over
number counts alone, but little improvement for less dense ones.

In this paper, we combine the halo model reconstruction approach of
\citet{GunnarssonEtal2006}, \citet{JonssonEtal2010}, 
\citet{WongEtal2011} and others, with the
idea of calibrating to simulations from \citet{SuyuEtal2010}
Taking the $\kappax$ values from
\citet{HilbertEtal2009}'s ray-tracing calculations as the ``truth''
that we have to recover, we use the \MS mock galaxy catalogues first to
{\it calibrate} the reconstruction and account for unseen mass and
voids, and then to {\it test} the accuracy of the line of sight mass
reconstruction under this assumption. Probabilistically assigning mass
and redshift to every observed galaxy in a given observed field, we
generate Monte Carlo sample line of sight mass distributions, and so
construct the probability distribution function (PDF) $\Pr(\kappax|\data)$ for that field.
We then emulate
the combination of many such PDFs to quantify the residual bias that
would be translated to the global (including cosmological) parameters, 
\phil{if left unaccounted for.}
In doing so, we aim to answer the following questions:

@@

In strong lens systems where the source is time variable, the images do not 
vary simultaneously; the optical path length for each image is different due
to relativistic and geometric effects. The difference in optical path length
causes a time-delay between the light curves of each image.
\citet{FalcoEtal1985} showed that the presence of additional matter (in the
form of a mass-sheet) along the line of sight has no observable effect except
to rescale the time-delays and magnifications. As such, it is necessary to include $\kappax$ in
the lens modelling if cosmological parameters are to be estimated accurately
and precisely from observed time delays \citep{SuyuEtal2010}.

If there is external convergence present that is not included in the
lens modelling, then the time delay distance -- inferred assuming $\kappax
= 0$ -- will be $(1-\kappax)$ more than the true value of the time-delay distance $\Dt$:
\be 
\label{eq:MassSheet:Dtbias}
\Dt^{\rm{true}}=\frac{\Dt^{{\kappax = 0}}}{1-\kappax}.
\ee

@@

Dynamical observations of the lens galaxy
can help break the mass-sheet degeneracy by providing an additional
estimate of the lens's mass
\citep[e.g.,][]{KoopmansEtal2003,Koopmans2004,SuyuEtal2010}. This constraint 
is useful in excluding the high convergence tails that are often present in 
PDFs for convergence - we do not include dynamical constraints in our analysis.

In this paper we restrict ourselves to reconstructing
the line of sight portion of the mass distribution in any given field, {\it our lines of sight do not include a strong lens} which allows us to work in the weak lensing regime.

The impact of weak-lensing perturbations on strong lensing lines-of-sight is
postponed to further work.

@@

\citet{HilbertEtal2009} calculated the lensing convergence using the
``lightcone'' dark matter only  output of the \MS, providing 1.7 degree
square  mock sky maps of this quantity. In this process, the dark matter
density from the simulation was projected onto a set of lens planes at
discrete redshifts, adaptively gridded and smoothed, and then the second
derivatives of the  lensing potential required for the components of the
magnification matrix were computed using a multiple lens plane ray tracing
algorithm. The first order approximation to this algorithm \citep[equation 17
of][]{HilbertEtal2009} gives the total convergence at a given sky
position~$\bmath{\theta}$ as the simple summation of the surface density in
each lens plane, weighted by the inverse critical surface density $\Sigma_{\rm
cr}$ for that plane's redshift:
\be
\kappax(\bmath{\theta}) = \sum_i \frac{\Sigma_i(\bmath{\theta})}
                                     {\Sigma_{\rm cr}(z_i,\zs)}.
\ee
\citet{HilbertEtal2009} found that this approximation is accurate to a percent level
in $\kappax$, even on 30 arcsecond scales. This accuracy sufficient for our analysis, but 
the approximation may need re-visiting in future work.

@@

\new{The adaptive smoothing of the matter in the \MS results in a spatial resolution of $\sim 5\,\text{kpc}$ comoving in the densest regions, and $\sim10\,\text{kpc}$ on average. This resolution is sufficient to capture most of the variance in the convergence \citep{TakahashiEtal2011}. Moreover, we seek to estimate the component of the convergence that is smooth on the angular scales of typical strong galaxy-scale lens systems ($\sim 1\,\text{arcsec}$). Any variations of the external convergence on smaller scales can be directly recovered from the resulting image distortions in the lens modelling.}

@@

We construct mock photometric galaxy catalogues by selecting all \MS galaxies
within a circular aperture of a given angular radius centred on each randomly
selected line of sight position, with apparent $i$-band magnitude brighter
than a given limit. The parent catalogue, from \citet{HilbertEtal2011},
contains galaxy positions and magnitudes, all of which include lensing effects
(deflections and magnifications). We also have access to the underlying galaxy
halo masses and redshifts, which we will use to calibrate the reconstruction,
and also to explore any sources of bias and scatter. Since the ``true''
convergence is that of only the dark matter, we focus on reconstructing the
dark matter halos alone. Our model for doing this is outlined in the next
section.

@@

Cosmological dark matter simulations have shown that dark matter
halos are reasonably well-approximated by NFW profiles 
\citep{NFW}. We assume each halo to have a spherical mass distribution with 
density given by
\be\label{eq:rhonfw}
\rho(r) = 
\frac{\rho_0}{\left(r/r_{s}\right)
\left(1+r/r_{s}\right)^{2}}.
\ee
Here, $r_{s}$ is a characteristic scale 
radius of the cluster, representing the point where
the density slope transitions from $r^{-1}$ to $r^{-3}$. This radius is
related to the virial radius of the halo by $r_{s}~=~r_{200}/c$, where $c$ is
the concentration parameter, which can be estimated from the halo's mass,
using a mass--concentration relation. Typically more massive halos are less
concentrated, but there is some scatter; we use the relation of
\citet{Neto2007} to estimate $c$ from the mass enclosed within $r_{200}$,
which we denote as $M_{200}$. We find our results do not change if the
\citet{MaccioEtal2008} mass--concentration relation is used instead.

@@

Many studies have shown that galaxies have total mass density profiles that
are approximately isothermal in their inner regions
\citep[\eg][]{AugerEtal2010} due to the more concentrated stellar mass
component.
This has been confirmed by galaxy-galaxy weak lensing measurements 
\citep[\eg][]{Mandelbaum,GavazziEtal2007}. We note that the two halo term that
dominates the galaxy-galaxy lensing signal at large radii is explicitly taken
into account in our model, where every galaxy is assigned a dark matter halo.
At large radii an NFW-like profile is probably appropriate, but may not be
correct when a halo is very close to the line of sight.

@@

To estimate a halo's contribution to $\kappah$ we must first know the
halo's mass \new{and position. The stellar mass
and host halo are not necessarily concentric, especially for central
galaxies of giant clusters, but our reconstruction assumes the dark matter halo is centred on
the visible galaxy.} For the lines of sight of the \MS, halo masses are
known, but for  real lines of sight the halo masses have to be \infered
from whatever data are available. 
In this work we investigate the use of the empirical stellar mass --
halo mass relation for this purpose. The primary advantage of this
approach is that it is a simple, ``one size fits all'' relation that can
be applied regardless of the stellar mass observed. We discuss
additional observables that could be used to improve the precision of
the halo mass inference in \Sref{sec:discuss} below.

@@

The solution we present here is to calibrate the relationship between the halo
convergence $\kappah$ and the overall convergence $\kappax$ (as would be
calculated by ray tracing through the full density field)  using the
simulations themselves. The halo modelling procedure outlined in
\Sref{sec:model:halos} allows  us to estimate  the convergence due to halos
given observations of  mass, redshift and position,
$\pr(\kappah|\data)$, but what we are interested in is the total
convergence along the line of sight $\pr(\kappax|\data)$. We can obtain
this by considering the expression: 
\begin{equation} 
\Pr(\kappax|\data) = 
   \int \dee\kappah  \Pr(\kappax|\kappah,\data) \Pr(\kappah|\data)
   \label{eq:kappaconv}
\end{equation} 
The first term in the integrand relates the convergence due to model halos,
$\kappah$, to the true convergence, $\kappax$. This conditional distribution
can be constructed from the \MS catalogues, by computing
$\kappah$ from the true halo masses and redshifts 
along each selected line of sight, and then
accumulating $\kappah$, $\kappax$ pairs
for a large number of such sightlines. 
For any given line of sight we can then estimate
$\pr(\kappah|\data)$ as described above, and then multiply it by
$\Pr(\kappax|\kappah)$ and integrate out the intermediate parameter
$\kappah$.

@@

We now put the reconstruction \proceedure outlined above into practice, first
inferring $\kappax$ from $\kappah$ in the case of hypothetical galaxy
catalogues with noiseless redshifts and halo masses. The reason for doing this
is to check the validity of the calibration process, and then also to 
investigate the primary sources of the convergence. It will also provide us
with a measure of the intrinsic uncertainty introduced by our assumptions of
all illuminated halos being spherically-symmetric truncated NFW halos following the
Neto et al concentration--mass relation \new{and of visible galaxies being concentric with their host halos}.

@@

\Fref{fig:jointkh-k} shows the conditional distribution of $\kappax$ given 
$\kappah$, derived from $10^5$ randomly selected \MS lines of sight
and assuming noise-free halo masses and redshifts. 
\phil{In this case, $\pr(\kappah|\data)$ is almost a delta function since the only uncertainty is halo
concentration, which has minimal effect on $\pr(\kappah|\data)$.  
}
% In this case, the PDF : we can read
% off  $\pr(\kappax|\data)$ just by taking a thin vertical slice through this 
% distribution. 
% Note that the joint PDF plotted in this figure is the product
% of  the conditional distribution that we need for calibration, and also the
% prior PDF for $\kappah$ that we need to ensure that the global distribution of
% $\kappah$ values is defaulted to in the case of poor observational data. 
At fixed $\kappah$ we find in \Fref{fig:jointkh-k} that the scatter in 
$\kappax$
grows with $\kappah$; our reconstruction is better at reproducing under-dense
lines of sight than over-dense lines of sight. The effect of ignoring voids
and the smooth mass component is evident in this plot: $\kappah$ is
significantly higher than $\kappax$, at any given $\kappax$ value. 
\phil{Conditioning on $\kappah$ gives a PDF for
$\kappax$ centred at the correct place, by construction.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\includegraphics[width=\textwidth]{figs/where_is_the_kappa.eps}
\caption[magcut]{Which objects dominate the external convergence for a line of
sight? From left to right, the three figures show the cumulative contribution
to the total convergence from individual halos ($\kappah$) as a function of
1) distance from the line of sight, 2) magnitude and 3) redshift.   In each
panel, the solid line is the median cumulative contribution over a large
sample of lines of sight, while dashed and dot-dashed show the ranges enclosing 68
and 95 percent of the lines of sight respectively. The source is at redshift 1.4.}
\label{fig:where}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We first consider a purely random selection of lines of sight. With the purely
photometric reconstruction, there is no evidence of bias, even after combining
20 lines of sight. \tom{This result is a sanity check: if we had chosen lines
of sight with identical $\kappah^{\rm med}$ rather than $\kappah^{\rm
med}\pm0.003$, randomly selected lines of sight would have zero bias by
construction}. If lines of sight hosting time delay lenses are truely random, 
our method allows $\kappax$ to be reconstructed without inducing a bias (under
the basic assumptions that the universe is like the calibration simulation,
and the correct stellar-mass to halo-mass relation has been used).
